The Spatial and Temporal Domains of Modern Ecology

### 2017-12-19 paper/SI.tex
Revising main text to include information on observational type distributions, and updated figure 1 vignette to produce percentage of observations per bin by observations type
Finished revision to results in main text, moved trend analysis discussion mostly to SI

The following text provides additional information on this study's Methods and Results.  The full manuscript, data, and code are available on an accompanying \href{https://github.com/agroimpacts/ecoscales}{github repository}. \textbf{Note to editors and reviewers: this repository is currently private, but will be made public upon acceptance of manuscript. In the interim, the entire project folder can be downloaded from the following \href{https://www.dropbox.com/sh/1l9em6p0wyabwr3/AADhCpFvqjv6y7VbL5tgB3Iqa?dl=0}{link}.}

@@

\subsubsection*{Trends in observational methods and scales by year}
\vspace{-10pt}
Figures \ref{type_by_yr} and \ref{sc_by_yr} respectively show how the frequency of ecological observing methods and the average scale in each dimensions varies by publication year (2004-2014).

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\textwidth]{../vignettes/figures/figS1.png}
\vspace{-20 pt}
\caption{Histograms of the resolution, extent, interval, and duration of observational scales, shown for all observational methods (top row) and for the four main observational methods (rows 2-5). Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval. Bar widths for interval and duration indicate differences in scale between x-axis labels. }
\label{obshists}
\end{figure}

>>>>

The following text provides additional information on this study's Methods and Results.  The full manuscript, data, and code for all calculations are available on an accompanying \href{https://github.com/agroimpacts/ecoscales}{github repository}. \textbf{Note to editors and reviewers: this repository is currently private, but will be made public upon acceptance of manuscript. In the interim, the entire project folder can be downloaded from the following \href{https://www.dropbox.com/sh/1l9em6p0wyabwr3/AADhCpFvqjv6y7VbL5tgB3Iqa?dl=0}{link}.}

@@

\subsubsection*{Trends in observational methods and scales by year}
\vspace{-10pt}
To evaluate the potential impact that excluding studies from 2015-2017 would have on our findings, we analyzed the trends in 1) ecological observing methods, and 2) typical scales of ecological observations over the 10 year period.

We used linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period. The results (Fig. \ref{type_by_yr}) suggest an increase (by 1.3\% per year) in the use of remote sensing and a corresponding decline in field methods (by the same percentage) during the 2004-2014 time period, although neither trend is statistically significant (remote observation regression, p$<$0.12; field observation regression, p$<$0.18).  Automated sensing methods showed no visible trend.

If the trend towards increasing use of remote sensing between 2004-2014 was real and not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average.

Of course, we cannot attach confidence to a trend in observational extent extrapolated from a non-significant trend in observational type. However, some additional evidence in support of this possible trend lies within the extent values themselves, which appeared to increase by 0.25 orders of magnitude per year (Fig. \ref{sc_by_yr}) between 2004-2014 (R$^2$ = 0.25, p$<$0.07). This somewhat clearer (albeit still not significant at the customary p$<$0.05 threshold) trend suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).

The trend calculations described above can be found in the ``Additional Analyses" vignette in the accompanying \href{https://github.com/agroimpacts/ecoscales}{R package/code repository}.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\textwidth]{../vignettes/figures/figS1.png}
\vspace{-5 pt}
\caption{Histograms of the resolution, extent, interval, and duration of observational scales, shown for all observational methods (top row) and for the four main observational methods (rows 2-5). Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval. Bar widths for interval and duration indicate differences in scale between x-axis labels. }
\label{obshists}
\end{figure}

### 2017-12-19 paper/main.tex
Finished revision to results in main text, moved trend analysis discussion mostly to SI

\noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
For our final analysis of observational domains, we evaluated the differences between the scales represented by the extent and duration dimensions and the scales that ecological observations actually cover. To make this assessment, we log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, yielding the magnitude of difference between each pair of dimensions for each observation. We plotted these values (summarized in box plots) against their corresponding extent/duration values to evaluate whether these differences varied with scale (Fig. 3). Extent was on average 5.6 orders of magnitude larger than actual extent, and this difference increased with extent, reaching a maximum of 8 orders of magnitude between 100 million and 1 billion ha of extent (Fig. 3A). The difference fell to 3 orders of magnitude at 10 billion ha, a domain that was covered by $<$2\% of observations that were primarily collected with remote sensing. Remote observations had the smallest mean difference magnitude (1.9), whereas differences for the other three methods averaged at least 5.7 orders of magnitude larger (Fig. S3 in SI).

The magnitude of difference between the duration and actual duration of observations was somewhat smaller, averaging 3.4 across all observations, ranging from just under 2 for the shortest durations (hour-day) to over 4 for observations lasting 1 decade to 1 century (Fig. 3B). As with extent, the difference fell substantially for the longest durations (century to 10,000 years), as these domains were covered by paleo-reconstructions (Fig. S3 in SI), which have effectively no difference between duration and actual duration because the coring techniques they are typically based on capture continuous temporal records. The mean difference magnitudes for the other three observing methods ranged from just over 3 for field and automated sensing methods to nearly 6 for remote observations (Fig. S3 in SI).

@@

Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates.

Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance. Automated sensing methods showed no trend over time (SI).

Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average (see SI for details of calculation). Further evidence for this trend lies within the extent values themselves, which increased 0.25 orders of magnitude per year between 2004-2014 (R$^2$ = 0.25, p$<$0.07). This somewhat clearer trend also suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).

>>>>

\noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
For our final analysis of observational domains, we evaluated the differences between the scales represented by the extent and duration dimensions and the scales that ecological observations actually cover. To make this assessment, we log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, yielding the magnitude of difference between each pair of dimensions for each observation. We plotted these values (summarized in box plots) against their corresponding extent/duration values to evaluate whether these differences varied with scale (Fig. 3). Extent was on average 5.6 orders of magnitude larger than actual extent, and this difference increased with extent, reaching a maximum of 8 orders of magnitude between 100 million and 1 billion ha of extent (Fig. 3A). The difference fell to 3 orders of magnitude at 10 billion ha, a domain that was covered by the $<$2\% of observations that were primarily collected with remote sensing. Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (see Fig. S3 in SI for method-specific plots).

The magnitude of difference between the duration and actual duration of observations was somewhat smaller, averaging 3.4 across all observations, ranging from just under 2 for the shortest durations (hour-day) to over 4 for observations lasting 1 decade to 1 century (Fig. 3B). As with extent, the difference fell substantially for the longest durations (century to 10,000 years), as these domains were covered by paleo-reconstructions (Fig. S3 in SI), which have effectively no difference between duration and actual duration because coring techniques capture continuous temporal records. The mean difference magnitudes for the other three observing methods ranged from just over 3 (field observations and automated sensing) to nearly 6 (remote observations).

@@

The relatively small sample size of our study may also bias these findings. Although we used a randomized title draw to obtain a representative sample, we reviewed just 0.8\% of papers published during the 10 year period. Our sample may therefore under- or over-represent observational coverage in certain scale domains, particularly for analyses broken down by observational method. The case where this is of most concern is for paleo-reconstructions, the small sample size of which resulted in a likely over-estimate of the typical extent of such observations (e.g. Fig. 2B, C; however, the interval and duration findings are probably more representative).

Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates. To evaluate this possibility, we used linear regressions to examine whether the relative frequency of observing methods changed during the 10 year study period, and whether our findings regarding scale domains might be different if we included more recent studies. Although our sample size was too small to assign statistical significance, we found a possible positive trend in the use of remote observations and a corresponding decline in field observations over the 10 year period. If these trends were not spurious, they suggest that including studies from 2015-2017 would result in a somewhat larger relative sample of remote observations, which would in turn increase slightly the mean extent of ecological observations (see SI for details of trend analyses).
